---
title: "WTO Timeseries API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{timeseries_database}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyverse)
library(wtor)
```

# Overview

wtor Timeseries API-related functions allow you to explore and retrieve all the time series in the excellent WTO database. 

The functions can be divided in two main groups:

* **Functions that allow you to retrieve the final data or that are related to this goal**. The main example being `get_timeseries_data()`
* **Accesory functions that retrieve metadata**. A very important one being `get_timeseries_available_indicators()`

## Caching

By default, requests to the Timeseries API are cached to disk.

# Getting started

First of all, set up your WTO API key if you haven't already. Sign up at [WTO API portal](https://apiportal.wto.org/) and retrieve your API key at your profile. Then, set environment variable WTO_R_API_KEY to the value of your key. Example: `usethis::edit_r_environ()` and then: `WTO_R_API_KEY=6bf2c780decc45c08e668f1d7d488534`. Lastly, restart your R session.

To test if the package has been correctly set up and working, retrieve a list of all the available indicators: `wtor::get_available_indicators()`.

# Trade-flows

The WTO Timeseries API includes a rich set of bilateral trade flows between members, under the category of "International Trade Statistics". To retrieve a list of all the available indicators in this category, execute `wtor::get_timeseries_available_indicators() |>  dplyr::filter(categoryCode == "ITS")`.

## Total trade flows of a member

Let us retrieve and plot the total value of merchandise exports reported by Albania, to the world economy, grouping by sector of exports:

```{r}

albania_exports_by_sector_df <- wtor::get_timeseries_data(code="ITS_MTV_AX", reporting_economies = "008") |> 
  left_join(
    wtor::get_products_sectors("SITC3") |> rename(productorsectorcode = code),
    by="productorsectorcode"
  ) |> 
  rename(sector = productorsectorcode) |> 
  select(year, value, reportingeconomycode, name) 

ggplot(data=albania_exports_by_sector_df,
       mapping=aes(
         x=as.integer(year),
         y=value,
         color=name)) +
  geom_line()

```

## Bi-lateral services trade flows

Let us now retrieve the bilateral services flows between two countries, France and Spain.

```{r}

# France's ISO 3166 country code is 250
# Spain's ISO 3166 country code is 724

spain_france_bilateral_flows_df <- wtor::get_timeseries_data(code="BAT_BV_X", reporting_economies = "724", partner_economies = "250") |> 
  left_join(
    wtor::get_products_sectors("BOP6") |> rename(productorsectorcode = code),
    by="productorsectorcode"
  ) |> 
  rename(sector = productorsectorcode) |> 
  select(year, value, reportingeconomycode, name) 

ggplot(data=spain_france_bilateral_flows_df,
       mapping=aes(
         x=as.integer(year),
         y=value,
         color=name)) +
  geom_line()

```


## Bilateral merchandise flows by MT codes

Let us retrieve the imports of a given reporting economy from another partner economies, by detailed MT code. We will then represent it as a scatter plot in which the X-axis will represent the value of American imports from Turkey, and the Y-axis will correspond to the value of the Turkish imports from the United States, for every MT sector.

```{r}

country1_display_name <- "USA"
country2_display_name <- "Turkey"

country1_code <- wtor::get_reporting_economies() |> filter(grepl("United States of America", name)) |> _$code
country2_code <- wtor::get_reporting_economies() |> filter(grepl("Tür", name)) |> _$code

bilateral_trade_df <- wtor::get_timeseries_data(code="HS_M_0020", reporting_economies = country1_code) |> filter(partnereconomycode == country2_code) |> 
  bind_rows(wtor::get_timeseries_data(code="HS_M_0020", reporting_economies = country2_code) |> 
              filter(partnereconomycode==country1_code)
  ) |> 
  left_join(
    wtor::get_reporting_economies() |> rename(reportingeconomycode = code),
    by="reportingeconomycode"
  ) |> 
  filter(stringr::str_length(productorsectorcode) == 2) |> 
  left_join(
    wtor::get_products_sectors("MT2") |> rename(productorsectorcode = code, sector_name = name) |> filter(stringr::str_length(productorsectorcode)==2),
    by="productorsectorcode"
  ) |> 
  select(name, sector_name, year, value) |> 
  pivot_wider(id_cols = c("sector_name", "year"), names_from="name", values_from="value") |> 
  mutate(year = as.integer(year)) |> 
  rename(country1=3,
         country2=4)

bilateral_trade_df[is.na(bilateral_trade_df)] <- 0

ggplot(data=bilateral_trade_df,
       mapping=aes(
         x=country1,
         y=country2,
         color=year
         )) +
  geom_point() +
  # scale_color_brewer(palette = "Set2") +
  scale_color_gradient(low = "red", high = "darkblue") +
  scale_x_continuous(labels = scales::number_format(accuracy=1, scale=1e-6, suffix="M", big.mark=".", decimal.mark=",")) +
  labs(x=country1_display_name,
       y=country2_display_name)

```
## Comparison of MFN tariff rates

Let us compare the simple average ad-valorem duty MFN rates of two given countries, by plotting the MFN rates they apply in a scatter plot, with every point representing a given MT sector average MFN rate.

```{r}

country1_display_name <- "Brazil"
country2_display_name <- "India"

country1_code <- wtor::get_partner_economies() |> filter(grepl(country1_display_name, name)) |> _$code
# country2_code <- wtor::get_partner_economies() |> filter(grepl(country2_display_name, name)) |> _$code

country2_code <- "356"

indicator_code <- "HS_A_0010"

mfn_rates_df <- bind_rows(
  wtor::get_timeseries_data(code=indicator_code, reporting_economies =  country1_code),
  wtor::get_timeseries_data(code=indicator_code, reporting_economies =  country2_code),
)  |> 
  left_join(
    wtor::get_reporting_economies() |> rename(reportingeconomycode = code),
    by="reportingeconomycode"
  ) |> 
  left_join(
    wtor::get_products_sectors() |> rename(productorsectorcode=code, sector_name=name),
    by="productorsectorcode"
  ) |> 
  filter(stringr::str_length(productorsectorcode) == 2) |> 
  select(year, name, productorsectorcode, sector_name, value) |> 
  # group_by(name) |> 
  # filter(year==max(year)) |> 
  pivot_wider(id_cols = c("year", "sector_name", "productorsectorcode"), names_from="name", values_from="value") |> 
  rename(
    country1_rate = 4,
    country2_rate = 5
  )

ggplot(
  data=mfn_rates_df,
  mapping=aes(
    x=country1_rate,
    y=country2_rate,
    color=as.integer(year),
    alpha=0.5
  )
) +
  geom_point() +
  scale_color_gradient(low = "red", high = "darkblue") +
  labs(x=country1_display_name,
       y=country2_display_name)

```

## Average MFN rates by year, for all reporting countries

## A map of MFN tariff rates by countries

# Bi-lateral tariff rates

```{r}

country1_display_name <- "United States of America"
country2_display_name <- "China"

country1_code <- wtor::get_reporting_economies() |> filter(grepl(country1_display_name, name)) |> _$code
country2_code <- wtor::get_reporting_economies() |> filter(grepl(country2_display_name, name)) |> _$code

country1_code <- "840"
country2_code <- "156"

indicator_code <- "HS_P_0070"

bilateral_rates_df <- bind_rows(
  wtor::get_timeseries_data(code=indicator_code, reporting_economies =  country1_code),
  wtor::get_timeseries_data(code=indicator_code, reporting_economies =  country2_code),
) |> 
  left_join(
    wtor::get_reporting_economies() |> rename(reportingeconomycode = code),
    by="reportingeconomycode"
  ) |> 
  left_join(
    wtor::get_products_sectors() |> rename(productorsectorcode=code, sector_name=name),
    by="productorsectorcode"
  ) |> 
  # filter(stringr::str_length(productorsectorcode) == 2) |> 
  select(year, name, productorsectorcode, value) |> 
  # group_by(name) |> 
  # filter(year==max(year)) |> 
  pivot_wider(
    id_cols = c("year", "productorsectorcode"),
    names_from="name", 
    values_from="value"
    ) |> 
  rename(
    country1_rate = 2,
    country2_rate = 3
  )


ggplot(
  data=mfn_rates_df,
  mapping=aes(
    x=country1_rate,
    y=country2_rate,
    color=as.integer(year),
    alpha=0.5
  )
) +
  geom_point() +
  scale_color_gradient(low = "red", high = "darkblue")


```



